---
title: "Cyclophilin D Knockout Protects the Mouse Kidney against Cyclosporine A-Induced Oxidative Stress and DNA Hypomethylation"
author: ["<span class=\"presenter\">Bruce Kirkpatrick<sup>1,3</sup></span>, Jost Klawitter<sup>1</sup>, Alexander Pennington<sup>1</sup>, Joshua M. Thurman<sup>2</sup>, Nidhi Kotecha<sup>1</sup>, Uwe Christians<sup>1</sup>, Jelena Klawitter<sup>1</sup><ol class=\"affiliations\">University of Colorado Anschutz Medical Campus&ensp;<li>Department of Anesthesiology</li>&ensp;<li>Division of Nephrology and Hypertension</li>&ensp;<li>Medical Scientist Training Program</li></ol>"]
output:
  drposter::drposter_poster:
    fill_page: FALSE
    self_contained: FALSE
csl: american-chemical-society.csl
references:
- id: ref1
  title: Cyclophilin D-dependent mitochondrial permeability transition regulates some necrotic but not apoptotic cell death
  container-title: Nature
  issued: 
    year: 2005
- id: ref2
  title: Cyclophilin D is a component of mitochondrial permeability transition and mediates neuronal cell death after focal cerebral ischemia
  container-title: PNAS
  issued: 
    year: 2005  
- id: ref3
  title: Oxidative stress caused by mitochondrial calcium overload
  container-title: Ann NY Acad Sci
  issued: 
    year: 2010      
- id: ref4
  title: 'Oxidative and nitrosative stress in kidney disease: a case for cyclosporine A'
  container-title: J Nephrol
  issued: 
    year: 2005
- id: ref5
  title: 'Mechanisms of cyclosporine-induced renal cell apoptosis: a systematic review'
  container-title: Amer J Nephrol
  issued: 
    year: 2013
- id: ref6
  title: 'Urine metabolites reflect time-dependent effects of cyclosporine and sirolimus on rat kidney function'
  container-title: Chem Res Toxicol
  issued: 
    year: 2009
- id: ref7
  title: 'Association of immunosuppressant-induced protein changes in the rat kidney with changes in urine metabolite patterns: a proteo-metabonomic study'
  container-title: J Proteome Res
  issued: 
    year: 2010 
- id: ref8
  title: 'Low-salt diet and cyclosporine nephrotoxicity: changes in kidney cell metabolism'
  container-title: J Proteome Res
  issued: 
    year: 2012  
- id: ref9
  title: 'Metabolic profiles in urine reflect nephrotoxicity of sirolimus and cyclosporine following rat kidney transplantation'
  container-title: Nephron Exp Nephrol
  issued: 
    year: 2009   
- id: ref10
  title: 'Differential effects of cyclosporine and tacrolimus on arterial function'
  container-title: Transpl Int
  issued: 
    year: 2011   
- id: ref11
  title: 'Effect of tacrolimus and cyclosporine on renal microcirculation and nitric oxide production'
  container-title: Transpl Proc
  issued: 
    year: 2004       
- id: ref12
  title: 'Nuclear association of a T-cell transcription factor blocked by FK-506 and cyclosporin A'
  container-title: Nature
  issued: 
    year: 1991
- id: ref13
  title: 'Calcineurin is a common target of cyclophilin-cyclosporin A and FKBP-FK506 complexes'
  container-title: Cell
  issued: 
    year: 1991
- id: ref14
  title: 'Exploring treatment options in renal transplantation: the problems of chronic allograft dysfunction and drug-related nephrotoxicity'
  container-title: Transplantation
  issued: 
    year: 2001
- id: ref15
  title: 'Long-term renal allograft survival: have we made significant progress or is it time to rethink our analytic and therapeutic strategies?'
  container-title: Am J Transpl
  issued: 
    year: 2004 
- id: ref16
  title: 'Twenty-third official adult heart transplantation report'
  container-title: J Heart Lung Transpl
  issued: 
    year: 2006  
- id: ref17
  title: 'Postconditioning inhibits mitochondrial permeability transition'
  container-title: Circulation
  issued: 
    year: 2005  
- id: ref18
  title: 'Inhibiting mitochondrial permeability transition pore opening: a new paradigm for myocardial preconditioning?'
  container-title: Cardiovasc Res
  issued: 
    year: 2002   
- id: ref19
  title: 'Postconditioning inhibits mitochondrial permeability transition'
  container-title: Future Cardiol
  issued: 
    year: 2005   
- id: ref20
  title: 'Inhibition of anoxia-induced injury in heart myocytes by cyclosporin A'
  container-title: J Mol Cell Cardio
  issued: 
    year: 1991       
- id: ref21
  title: 'Cyclosporine A prevents apoptosis- related mitochondrial dysfunction after neonatal cardioplegic arrest'
  container-title: J Thorac Cardiovasc Surg
  issued: 
    year: 2008
- id: ref22
  title: 'Loss of cyclophilin D reveals a critical role for mitochondrial permeability transition in cell death'
  container-title: Nature
  issued: 
    year: 2005
- id: ref23
  title: 'Cyclophilin D gene ablation protects mice from ischemic renal injury'
  container-title: Am J Phys Renal
  issued: 
    year: 2009
- id: ref24
  title: 'Knockdown of Cyclophilin D Gene by RNAi Protects Rat from Ischemia/ Reperfusion-Induced Renal Injury'
  container-title: Kidney & Blood P Res
  issued: 
    year: 2004 
- id: ref25
  title: 'Cyclophilin D and the mitochondrial permeability transition in kidney proximal tubules after hypoxic and ischemic injury'
  container-title: Am J Phys Renal
  issued: 
    year: 2011
- id: ref26
  title: 'Physiologic functions of cyclophilin D and the mitochondrial permeability transition pore'
  container-title: Circ J
  issued: 
    year: 2013
- id: ref27
  title: 'Cyclophilin D controls mitochondrial pore- dependent Ca(2+) exchange, metabolic flexibility, and propensity for heart failure in mice'
  container-title:  J Clin Invest
  issued: 
    year: 2010
- id: ref28
  title: 'CypD(-/-) hearts have altered levels of proteins involved in Krebs cycle, branch chain amino acid degradation and pyruvate metabolism'
  container-title:  J Mol Cell Cardio
  issued: 
    year: 2013                                
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)  # Hide code by default. Can override chunk-by-chunk
```


# {.col-2 style="grid-row: span 4;"}
<!-- Span the first column for four rows, that way we can capture the usage block, two rows of template examples, and refs/footer information -->

## Abstract 

We treated wild-type and mice lacking cyclophilin (CypD; _Ppif_ -/-) with cyclosporin A (CsA) to test the role and contribution of CypD to widely described CsA-induced renal toxicity and oxidative stress. Our results showed an increase in several known uremic toxins as well as oxidative stress markers prostaglandin F2&alpha; and 8-isoprostane in CsA-treated WT, but not in _Ppif_ -/- animals. Similarly, a decline in S-adenosylmethionine and the resulting methylation potential indicative of DNA hypomethylation was observed only in CsA-treated WT mice. This confirms previous reports of protective effects of CypD deletion on the mouse kidney mediated through a stronger resistance of these animals to oxidative stress and DNA methylation damage. However, negative effects of CsA on the glycolysis and overall energy metabolism in _Ppif_ -/- mice also indicated that additional, CypD-independent pathways are involved in the toxic effects of CsA on the kidney. 


## Background

<div class="fullwidth"> Mitochondrial dysfunction and oxidative stress have been implicated in CsA-induced nephrotoxicity. Controversial reports have suggested that CypD deletion may or may not be protective in the protection of cells against oxidative stress-induced cell death. </div>

### Cyclophilin D
CypD (encoded by the _Ppif_ gene) is an essential structural component of the mitochondrial permeability transition pore (MPTP) and as such involved in its opening and subsequent regulation of cell death.[@ref1;@ref2] Despite years of research, the mechanism of CsA nephrotoxicity remains poorly understood, however, with several potential mechanisms identified: oxidative stress,[@ref3;@ref4] apoptosis,[@ref5] metabolic changes [@ref6;@ref7;@ref8;@ref9] and an increase of vascular resistance resulting in decreased renal blood flow.[@ref10;@ref11]

### Cyclosporine
CSA blocks the phosphatase activity of calcineurin after binding to its cellular co-effectors, the cyclophilins. [@ref12;@ref13] CsA is responsible for a significant improvement of short-term survival of transplant organs. [@ref14] However, long-term graft survival in renal transplant recipients has only seen a marginal improvement. [@ref15;@ref16] As of today, approximately 50% of renal allograft transplant patients still lose their transplant after 12 years, among other reasons due to immunosuppressant-induced nephrotoxicity. 

### 
<div class="fullwidth">CsA has been shown to protect against ischemia/reperfusion (I/R) organ injury, [@ref17;@ref18;@ref19;@ref20;@ref21] an observation that most likely arises from CsA’s ability to inhibit CypD. Mice that lack CypD have enhanced capacity to retain mitochondrial calcium, show delayed opening of the MPTP when calcium overload is present, and have smaller infarctions and less damage after prolonged I/R.[@ref22;@ref23;@ref24;@ref1;@ref25] However, _Ppif_ -/- mice are more susceptible to heart failure initiated by different stimuli, such as exercise-induced hypertrophy. [@ref26;@ref27] Mechanistically, CypD ablation is associated with elevated levels of mitochondrial matrix Ca2+ that in turn leads to increased glucose oxidation relative to fatty acids. This metabolic switch limits the ability of organs, including kidney, to adapt during stress,[@ref26;@ref27;@ref28] and is similar to the effects we have observed after CsA treatment.[@ref6;@ref8;@ref9]</div>

## Hypothesis 

<div class="fullwidth">As one of the major potential molecular targets of CsA nephrotoxicity is missing, we hypothesized that mice with CypD deficiency will show a better outcome following CsA treatment with regard to the changes inflammatory response, less generation of oxidative stress and a higher renal energy capacity as compared to normal mice treated with CsA.</div>

## Methods

#### Mouse experiments

Eight-week old male and female mice (WT B6.129SF2 and _Ppif_ -/-) were given either skim milk or treated by oral gavage with CsA at (a) 25 mg/kg/day and (b) 50 mg/kg/day for 28 days. Animals were weighed weekly and blood glucose was monitored at the same time via tail vein prick blood collection. Urine and plasma collections occurred biweekly. At the end of treatment, the mice were sacrificed and blood, urine, and tissues were collected.

#### Kidney histology

Kidney tissue samples were fixed and embedded before staining in eosin solution. Stain intensity was optimized in ammonia water and sections were dehydrated in xylene. Images of the kidney sections were captured and 25 high-powered fields were examined in the tubulo-interstitium of the cortex and outer medulla of each section. The kidneys were examined for inflammation, epithelial necrosis, loss of brush border, and tubular dilatation.

#### Analyte quantification

Oxidative stress markers, creatinine, CsA, and high-energy phosphate metabolite concentrations were determined in extracted blood, plasma, urine, and snap-frozen tissue samples with LC-MS/MS. Corresponding deuterated internal standards were used for quantification.


#### Metabolomic analysis

Tissue samples were homogenized and subjected to a modified semi-quantitative targeted metabolomics. Positive identification of the metabolites of interest was performed through injection of pure compound standards. Some proteins were further analyzed by Western blot.


# {.col-2}

## Results

### Physiological and Histological Changes {.fullwidth}
 
Neither phenotype nor treatment with CsA for 28 days impacted animal weight gain or blood glucose levels. Inflammation was not observed in the glomerular or tubulointerstitial compartments of either strain of mice. Treatment with CsA was also not associated with tubular damage in either strain of mice. Vascular congestion was seen in the glomeruli of some _Ppif_ -/- mice but was seen in both control and CsA-treated mice.

### Creatinine & CsA

No changes in CsA blood concentration were observed between the mice (either sex-related or strain-based)  and plasma creatinine levels increased in all CsA-treated animals, but reached significance only in WT animals treated with 50 mg/kg/day CsA.

### Export

* View and "print as PDF" from Chrome[@firefox-footnote]  <!-- Note: the footnote used to be a bibtex "note," but the default ACS template doesn't include that field -->
* Be sure to save a PDF (and possibly html with `self_contained: true`) to archive your project at the end, in case there are changes in pandoc, rmarkdown, etc.
* You can also render the poster in other formats, such as a `github_document` or `revealjs::revealjs_presentation`


## Reproducible research

### Directly include plots

```{r, echo=FALSE}
plot(cars)
```

### Directly include stats

```{r cars_demo}
summary(cars)
```



# {.col-2}

## Customizable themes

Note the features for theming.  If you had a special class attached to the .level1 or .slides/theme, you could use a general descendent selector to automatically get theming support, then break up these details into separate theme files.

# {.col-4}

## Default theme

This is an example of the default theme.

# {.col-4 .theme-minimalist}

## Minimalist theme

This is an example of the minimalist theme.


# {.col-2}


## See also

* tikzposter latex template, and its example themes for inspiration <https://bitbucket.org/surmann/tikzposter/downloads/>
* [Other R markdown templates](https://gist.github.com/Pakillo/4854e5d760351206084f6be8abe476b2) with their advantages/disadvantages (pdf compatibility, consistent syntax with flexdashboard, etc.)

# {.col-1 .theme-minimalist} 

## References {style="--header-color:Black; --other-color-for-h2-gradient:Gray;"}
<div id="refs" class="references"></div>
